---
title: "Data Viz"
author: "Clarissa Boyajian"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# General use
library(tidyverse)
library(here)
library(janitor)

# Specifically for dataviz:
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

library(sf)
```

## Get the lizards data
```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv:"))
```

Make a histogram of lizard weights
**Can include line calling out what data and aesthetics to use in the graph type () instead of ggplot() - helpful if you're layering different plots with different data on same image**
```{r}
ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(fill = "orange", 
                 #fill is color of shapes
                 color = "purple", 
                 #color is for lines and points
                 size = 0.25, 
                 #change the border line
                 linetype = "dotted")
```

Make a scatter plot of total length on x-axis, total weight on y-axis
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 22, 
             #different numbers 0-25 refer to different shapes (some point styles have fills and colors)
             fill = "purple",
             color = "red",
             size = 2,
             alpha = 0.5) 
            #changes opacity
```

Bubble plot where the color of the points changes based on common_name adn the size of the points varies based on total length. (length, weight)
**This is a bad idea**

- based on variable, in `aes()`
- based on constant, outside of `aes()`
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_wrap(~common_name)
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_grid(sex ~ tail) 
  #rows, columns
```

Getting things in order
```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

#Can also do: `lizards %>% dplyr::count(common_name)`
#If wanting multiple summary values, makes more sense to do `group_by() %>% summarize()`
lizard_counts_cn_tail <- lizards %>% count(common_name, tail)
```

Plot
```{r}
ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, count), x = count)) +
#Can put y first and x second to have it flip the axis, good to do here because of long names
#`fct_reorder(variable to change to a factor, value to order it by)` - part of tidyverse
  geom_col(aes(fill = common_name), show.legend = FALSE) +
  labs(x = "Number",
       y = "Common name")

```

Try converting common name to an ordered factor outside of `ggplot`
```{r}
lizard_counts_factored <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, count))
```


## Axis scale breaks, limits, and labels

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme_light() +
  scale_x_continuous(breaks = c(0, 50, 500),
                     limits = c(0, 500), 
                     expand = c(0, 0)) +
  scale_y_continuous(breaks = c(0, 20, 40, 60, 80), 
                     limits = c(0, 80),
                     expand = c(0, 0))
```

Transform the date column to class Date, then find counts of observations by date:
```{r}
lizard_counts_date <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  group_by(date) %>% 
  summarize(count = n())
```


Make a line plot of date (x) and count (y)
```{r}
ggplot(data = lizard_counts_date, aes(x = date, y = count)) +
  geom_line() +
  labs(x = "Date",
       y = "Number of lizards counted") +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%Y")
```

Adding color to scatter plot:
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight), show.legend = FALSE) +
  scale_color_gradientn(colors = c("orange", "red", "purple", "black"))
  #Can give it a vector of colors to manually create gradient

```

Manually creating step-wise colors schemes (doesn't do gradients anymore)
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight), show.legend = FALSE) +
  scale_color_stepsn(colors = c("green", "blue", "purple", "pink"),
                     breaks = c(10, 20, 30, 40))

```

## Update color scheme using a palette in paletteer:

Make a horizontal boxplot with common name on y, total length on x, with color changing based on common name
```{r}
lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

ggplot(data = lizards_fct, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "trekcolors::andorian")

```

## Fine tuning within theme
**need to know what object is stored that you want to change, what kind of element that is, what aesthetic you're changing**

Grid lines and background colors
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.background = element_rect(fill = "yellow", 
                                    color = "purple",
                                    size = 10),
    panel.grid.major.x = element_line(color = "red"),
    panel.grid.major = element_line(color = "blue"),
    #most specific level trumps other things within theme
    panel.grid.minor.y = element_line(color = "orange")
    #use `panel.grid = element.blank` to create no grid lines
  )

```

Window, axis text, labels
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.grid = element_blank(),
    plot.background = element_rect(fill = "cyan4"),
    axis.text.x = element_text(color = "orange"),
    axis.title = element_text(color = "yellow", size = 15)
  )

```



