---
title: 'EDS 221: Day 9'
author: "Clarissa Boyajian"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# General use packages:
library(tidyverse)
library(here)
library(janitor)

# Specifically for plots:
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

# And for another dataset we'll explore:
library(gapminder)

# Spatial (maps, maps, maps)
library(sf)
```


```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

## Use `ggrepel()`

Make a subset from lizards called ww_lizards, that only contains observations for "western whiptail" lizards from the site "sand"
```{r}
ww_lizards <- lizards %>% 
  filter(common_name == "western whiptail", site == "sand")

ggplot(data = ww_lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num), size = 3, max.overlaps = 20)
```

```{r}
#View(gapminder)

# Filter for only Europe, then create a scatterplot of gdpPercap (x) and lifeExp (y), label by country name

gapminder_eur <- gapminder %>% 
  filter(continent == "Europe", year == 2002)

ggplot(data = gapminder_eur, aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)

```

## gghighlight
```{r}
p <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point()

p + gghighlight(toe_num == 250, label_key = toe_num)
```

Create line plot of total length (x) vs weight (y) and highlight species with weight above 30
```{r}
q <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_line(aes(color = common_name)) + 
  gghighlight(max(weight) > 30)

q
```

## Use `patchwork()` for compound figures

```{r}
# can create a compound figure of graphs
# use & to add theme to all figures within compound figure
(p | q) / (q * p) / (q | q) &
  theme_minimal()
```

## A few new graph types

### Marginal plots

Marginal plot with rug
```{r}
whiptails <- lizards %>% 
  filter(common_name == "western whiptail") %>% 
  drop_na(total_length, weight)

ggplot(data = whiptails, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_rug()

```

Marginal plot with boxplots in margins
```{r}
ggplot(data = whiptails, aes(x = total_length, y = weight)) +
  geom_point(aes(color = sex), size = 2) +
  scale_color_manual(values = c("cyan4", "black", "goldenrod"),
                     name = "Sex:",
                     labels = c("female", "juvenile", "male")) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### Beeswarm plot

**can layer plots on top of each other**

**for categorical data, good idea to show actual data as beeswarm and then put summmary info on top**
```{r}
ggplot(data = whiptails, aes(x = sex, y = weight)) +
  geom_beeswarm() +
  geom_violin(fill = NA) +
  geom_boxplot(fill = NA)

```

### Headmap with `geomtile()`

Make new dataframe called lizard_counts, starting from lizards, with date converted to class Date, then count by year and common name the number of lizards observed
```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  mutate(year = lubridate::year(date)) %>% 
  group_by(year, common_name) %>% 
  summarize(count = n())
  # could use `count(year, common name)` instead of `group_by() %>% summarize(n())`
```

Make heatmap - fill color = amoutn of lizards each year
```{r}
ggplot(data = lizard_counts, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = count)) + 
  geom_text(aes(label = count), color = "white", size = 3) +
  scale_fill_viridis_c()
  # colorblind friendly color scheme
```

### Make a map

Map dominant vegetation types

```{r}
jornada_vegetation <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml")) %>% 
  select(Name) %>% 
  clean_names()

ggplot(data = jornada_vegetation) +
  geom_sf(aes(fill = name), color = NA) +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys") +
  labs(x = "Longitude",
       y = "Latitdue",
       fill = "Dominant vegetation:") +
  theme_classic()
```
















